Train Harder, Smarter:\\Using Graph Theory to Create Route Suggestions

### 2013-11-19 paper.tex
More 2d - 3d conversions...

The first weight is based on distance. The idea is that there is a theoretical,
perfect loop that covers exactly the amount of distance that you would like,
and the algorithm does its best to direct you along that loop. See
Figure~\ref{fig:ideal} for an example of what this might look like.

@@

In order to direct you along this loop, weights are generated based on
direction. We'll consider the method for finding the optimal direction,
$\theta$, briefly. First, consider how the weights are generated once
that direction is discovered. Basically, there is a weighted compass oriented
along $\theta$, with weights from 0.5 to 2. See Figure~\ref{fig:weights} to
see what we mean. As shown, the weights do not scale linearly. Instead, they
scale exponentially. This ensures that the weights are mostly centered around
the perpendicular, with growth accelerating farther from the perpendicular.
The idea is to have the weight be double in the ideal
direction, half moving exactly opposite the ideal direction, and unchanged
perpendicular to the ideal direction. Considering these three pieces of
information, we are looking for some $f : [0,2\pi] \to [0.5,2]$ as follows:
\begin{align*}
  2 = & f(\theta) \\
  1 = & f(\theta \pm \frac{\pi}{2}) \\
  \frac{1}{2} = & f(\theta \pm \pi)
\end{align*}
Let $\tau$ be the angle under consideration. We let
\[ \mbox{ang}(\tau, \theta) = |\theta - \tau \mbox{ mod } \pi|. \]
This allows us to define $f$ in terms of the angle between $\tau$ and
$\theta$. This function is necessary because $\theta$ is not necessarily
0. Note that this function only allows $\mbox{ang}(\tau, \theta)
\in [0,\pi]$. We accept this since we would like a compass symmetric
over the angle $\theta$, and we don't care if the angles are negative
or positive. However, the original function was defined in terms of the
value of $\tau$; instead, we are interested in a function based on the
angle between $\tau$ and $\theta$, which requires a slight redefinition:
\begin{align*}
  2 = & g(0) \\
  1 = & g(\frac{\pi}{2}) \\
  \frac{1}{2} = & g(\pi)
\end{align*}
We've already expressed the fact that we're interested in an exponential
function, and a base of two seems the obvious choice, leading to another
reduction:
\begin{align*}
  2 = 2^{h(0)} & \to & h(0) = 1 \\
  1 = 2^{h(\frac{\pi}{2})} & \to & h(\frac{\pi}{2}) = 0\\
  \frac{1}{2} = 2^{h(\pi)} & \to & h(\pi) = -1
\end{align*}
This yields two sets of three collinear points, which is easy enough to
map one to the other. Consider $h : [0,\pi] \to [-1,1]$ as follows:
\[ h(\mbox{ang}(\tau,\theta)) = 1 - \frac{2}{\pi}\mbox{ang}(\tau, \theta). \]
This satisfies the three points we gave, so we plug it in and calculate weight
accordingly:
\begin{equation}
  \mbox{WEIGHT} = 2^{1-\frac{2}{\pi}\mbox{ang}(\tau,\theta)}. \label{eq:weight}
\end{equation}

@@

The next step is to utilize these equations to calculate $\theta$. At any point
along the algorithm, we are at some $x_i$, and we have travelled a distance,
$d_i$. Figure~\ref{fig:algvals} should help to understand what exactly the
algorithm is attempting to achieve. The idea is this: if we have travelled some
distance $d_i$, then there is an $x_{\mbox{{\tiny ideal}}}$ on the ``ideal''
loop. The arc between $x_{\mbox{{\tiny ideal}}}$ and $x_0$ will be of length
$d_i$. The algorithm searches for a point $x_j$ such that the arc between
$x_{\mbox{{\tiny ideal}}}$ and $x_j$ will have the \emph{same} distance as the
line between $x_i$ and $x_j$. Fortunately, we have framed the equations for
calculating $x_j$ in terms of the distance travelled along the arc, and the
distance between $x_i$ and $x_j$ can easily be calculated using the pythagorean
theorem. Note that we find $x_j$ as follows:
\begin{align}
  x_j = \frac{d}{2\pi} ( & \cos( \frac{ 2 \pi }{ d } ( \min( d, d_i + d_j ) )
                                  + \phi ) \notag \\ & - \cos \phi ) + x_0
                                  \label{eq:xj} \\
  y_j = \frac{d}{2\pi} ( & \sin( \frac{ 2 \pi }{ d } ( \min( d, d_i + d_j ) )
                                  + \phi ) \notag \\ & - \sin \phi ) + y_0.
                                  \label{eq:yj}
\end{align}
To satisfy our constraint, we have
\[ d_j = \sqrt{ ( x_j(d_j) - x_i )^2 + ( x_j(d_j) - x_i )^2 }. \]

\subsubsection{Finding $d_j$}
Obviously, this is a rather difficult equation to truly solve, so we must only
heuristically find a solution. Unfortunately, even the heuristic solution is
a difficult one for a few reasons. Most problematically, the sine and cosine
functions have up to two stationary points on the interval we are considering
and there may not be a solution of $d_j$ that will satisfy the equality.
Nonetheless, we will do our best to mitigate these issues and apply a good
heuristic to find $d_j$. First, we rearrange our equation so that we can apply
a root-finding approximation algorithm to it:
\begin{equation}
  f(d_j) = \sqrt{(x_j(d_j) - x_i)^2 + (y_j(d_j) - y_i)^2} - d_j. \label{eq:dj}
\end{equation}
This simply allows the $d_j$ we are looking for to be the zero of the function.
The first exciting bit of news is that for most solutions of this equation with
$d_j < d$, it will work out that $f(0) > 0$ and $f(d-d_i) < 0$.
We know that $f(0) \geq 0$ by the
positivity of norms, and the square root of the sum of the squares is the
euclidian norm. We also can be certain that $f(d-d_i) < 0$ since
the norm is also unable to change by more than the diameter
of the ideal loop, or even less based on the proximity of $x_i$ to the center of
the loop. In situations where $f(d-d_i) > 0$, we notice that the distance to the
start of the loop is further away than the amount of distance we have left to
travel; this means that we should just route back to the start anyways. As such,
we have bounds inside of which we know that there exists a solution. We apply
Brent's method, a robust root approximation algorithm, to determine the solution
inside of this zone. The idea behind Brent's method is to combine the bisection
method, the secant method, and inverse quadratic interpolation. Readers
unfamiliar with any of these techniques should refer to CITATIONS. Some
pseudocode is given for Brent's Method in Algorithm~\ref{alg:brent}. The idea
is this:

@@

In general, the inverse quadratic interpolation can proceed very quickly, but we
add the bisection method to ensure convergence. In using interpolation, one of
the following conditions must be satisfied:
\begin{itemize}
\item Inverse Quadratic Interpolation was performed last time,
  \[ |s - b_k| < \left|\frac{b_{k-1}-b_{k-2}}{2}\right|, \]
  and
  \[|\delta| < |b_{k-1} - b_{k-2}| \]
\item The Bisection Method was used last time,
  \[|s - b_k| < \left|\frac{b_k-b_{k-1}}{2}\right|, \]
  and
  \[ |\delta| < |b_k - b_{k-1}| \]
\end{itemize}
The rest of the pseudocode serves only to ensure that variables are assigned
correctly. One notes that mflag indicates whether interpolation was used last
time or not.

\begin{algorithm}
  \caption{Using Brent's Method to find a zero of a function}
  \label{alg:brent}
  \begin{algorithmic}
    \Require $f(a_0)f(b_0) > 0$
    \Function{Brent}{$f : [a,b] \to \mathbb{R}$, $a_0$, $b_0$}
      \If{$|f(a_0)| < |f(b_0)|$}
        \Call{swap}{$a_0$,$b_0$}
      \EndIf
      \State $b_{k-1} \gets a_0$
      \State mflag $\gets$ True
      \While{$f(b_k) \not = 0$ {\bf and} $|b_k - a_k| > \epsilon$}
      \begin{align*}
        s \gets &\frac{a_kf(b_k)f(b_{k-1})}{(f(a_k)-f(b_k))(f(a_k)-f(b_{k-1}))} \\
        & + \frac{b_kf(a_k)f(b_{k-1})}{(f(b_k)-f(a_k))(f(b_k)-f(b_{k-1}))} \\
        & + \frac{b_{k-1}f(a_k)f(b_k)}{(f(b_{k-1})-f(a_k))(f(b_{k-1})-f(b_k))}
      \end{align*}
      \State {\bf if} $\displaystyle s \not \in \left[ \frac{3a_k+b_k}{4},b_k \right]$
             {\bf or}
             \State ~~~~(mflag is True {\bf and}
             \State ~~~~~~~~ ($\displaystyle |s-b_k| \geq \frac{|b_k-b_{k-1}|}{2}$
             \State ~~~~~~~~~~ {\bf or} $|b_k-b_{k-1}| < |\delta|$)) {\bf or}
             \State ~~~~(mflag is False {\bf and}
             \State ~~~~~~~~ ($\displaystyle |s-b_k| \geq \frac{|b_{k-1}-b_{k-2}|}{2}$
             \State ~~~~~~~~~~ {\bf or} $|b_{k-1}-b_{k-2}| < |\delta|$)) {\bf then}
             \State ~~~~ $\displaystyle s \gets \frac{a_k+b_k}{2}$
             \State ~~~~ mflag $\gets$ True
             \State {\bf else}
             \State ~~~~ mflag $\gets$ False
        \State $b_{k-2} \gets b_{k-1}$
        \State $b_{k-1} \gets b_k$
        \If{$f(a_k)f(s) < 0$}
          \State $b_k \gets s$
        \Else
          \State $a_k \gets s$
        \EndIf
        \If{$|f(a_k)| < |f(b_k)|$}
          \Call{swap}{$a_k$,$b_k$}
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{Weight for Elevation Gain}

Now that weights have been assigned based on distance and direction, we can
move on to worrying about elevation gain.

FILL IN THIS SECTION

@@

@@

@@

>>>>

The first weight is based on distance. The idea is that there is a theoretical,
perfect loop that covers exactly the amount of distance that you would like,
and the algorithm does its best to direct you along that loop. See
Figure~\ref{fig:ideal} for an example of what this might look like.In order to
direct you along this loop, weights are generated based on
direction. We'll consider the method for finding the optimal direction,
$\theta$.

@@

\subsubsection{Calculating the Weights}

First, consider how the weights are generated once
that direction is discovered. Basically, there is a weighted compass oriented
along $\theta$, with weights from 0.5 to 2. See Figure~\ref{fig:weights} to
see what we mean. As shown, the weights do not scale linearly. Instead, they
scale exponentially. This ensures that the weights are mostly centered around
the perpendicular, with growth accelerating farther from the perpendicular.
The idea is to have the weight be double in the ideal
direction, half moving exactly opposite the ideal direction, and unchanged
perpendicular to the ideal direction. Considering these three pieces of
information, we are looking for some $f : [0,2\pi] \to [0.5,2]$ as follows:
\begin{align*}
  2 = & f(\theta) \\
  1 = & f(\theta \pm \frac{\pi}{2}) \\
  \frac{1}{2} = & f(\theta \pm \pi)
\end{align*}
Let $\tau$ be the angle under consideration. We let
\[ \mbox{ang}(\tau, \theta) = |\theta - \tau \mbox{ mod } \pi|. \]
This allows us to define $f$ in terms of the angle between $\tau$ and
$\theta$. This function is necessary because $\theta$ is not necessarily
0. Note that this function only allows $\mbox{ang}(\tau, \theta)
\in [0,\pi]$. We accept this since we would like a compass symmetric
over the angle $\theta$, and we don't care if the angles are negative
or positive. However, the original function was defined in terms of the
value of $\tau$; instead, we are interested in a function based on the
angle between $\tau$ and $\theta$, which requires a slight redefinition:
\begin{align*}
  2 = & g(0) \\
  1 = & g(\frac{\pi}{2}) \\
  \frac{1}{2} = & g(\pi)
\end{align*}
We've already expressed the fact that we're interested in an exponential
function, and a base of two seems the obvious choice, leading to another
reduction:
\begin{align*}
  2 = 2^{h(0)} & \to & h(0) = 1 \\
  1 = 2^{h(\frac{\pi}{2})} & \to & h(\frac{\pi}{2}) = 0\\
  \frac{1}{2} = 2^{h(\pi)} & \to & h(\pi) = -1
\end{align*}
This yields two sets of three collinear points, which is easy enough to
map one to the other. Consider $h : [0,\pi] \to [-1,1]$ as follows:
\[ h(\mbox{ang}(\tau,\theta)) = 1 - \frac{2}{\pi}\mbox{ang}(\tau, \theta). \]
This satisfies the three points we gave, so we plug it in and calculate weight
accordingly:
\begin{equation}
  \mbox{WEIGHT} = 2^{1-\frac{2}{\pi}\mbox{ang}(\tau,\theta)}. \label{eq:weight}
\end{equation}

\subsubsection{Calculating the Ideal Loop}

@@

The next step is to utilize these equations to calculate $\theta$. At any point
along the algorithm, we are at some $x_i$, and we have travelled a distance,
$d_i$. Figure~\ref{fig:algvals} should help to understand what exactly the
algorithm is attempting to achieve. The idea is this: if we have travelled some
distance $d_i$, then there is an $x_{\mbox{{\tiny ideal}}}$ on the ``ideal''
loop. The arc between $x_{\mbox{{\tiny ideal}}}$ and $x_0$ will be of length
$d_i$. The algorithm searches for a point $x_j$ such that the arc between
$x_{\mbox{{\tiny ideal}}}$ and $x_j$ will have the \emph{same} distance as the
line between $x_i$ and $x_j$. Fortunately, we have framed the equations for
calculating $x_j$ in terms of the distance travelled along the arc, and the
distance between $x_i$ and $x_j$ can easily be calculated using the pythagorean
theorem. Note that we find $x_j$ as follows:
\begin{align}
  x_j = \frac{d}{2\pi} ( & \cos( \frac{ 2 \pi }{ d } ( \min( d, d_i + d_j ) )
                                  + \phi ) \notag \\ & - \cos \phi ) + x_0
                                  \label{eq:xj} \\
  y_j = \frac{d}{2\pi} ( & \sin( \frac{ 2 \pi }{ d } ( \min( d, d_i + d_j ) )
                                  + \phi ) \notag \\ & - \sin \phi ) + y_0.
                                  \label{eq:yj}
\end{align}
To satisfy our constraint, we have
\begin{equation}
d_j = \sqrt{ ( x_j(d_j) - x_i )^2 + ( x_j(d_j) - x_i )^2 }. \label{eq:dj}
\end{equation}

\subsubsection{Finding $d_j$}
Obviously, this is a rather difficult equation to truly solve, so we must only
heuristically find a solution. Unfortunately, even the heuristic solution is
a difficult one for a few reasons. Most problematically, the sine and cosine
functions have up to two stationary points on the interval we are considering
and there may not be a solution of $d_j$ that will satisfy the equality.
Nonetheless, we will do our best to mitigate these issues and apply a good
heuristic to find $d_j$. First, we rearrange our equation so that we can apply
a root-finding approximation algorithm to it:
\[ f(d_j) = \sqrt{(x_j(d_j) - x_i)^2 + (y_j(d_j) - y_i)^2} - d_j. \]
This simply allows the $d_j$ we are looking for to be the zero of the function.
The first exciting bit of news is that for most solutions of this equation with
$d_j < d$, it will work out that $f(0) > 0$ and $f(d-d_i) < 0$.
We know that $f(0) \geq 0$ by the
positivity of norms, and the square root of the sum of the squares is the
euclidian norm. We also can be certain that $f(d-d_i) < 0$ since
the norm is also unable to change by more than the diameter
of the ideal loop, or even less based on the proximity of $x_i$ to the center of
the loop. In situations where $f(d-d_i) > 0$, we notice that the distance to the
start of the loop is further away than the amount of distance we have left to
travel; this means that we should just route back to the start anyways. As such,
we have bounds inside of which we know that there exists a solution. We apply
Brent's method, a robust root approximation algorithm, to determine the solution
inside of this zone. The idea behind Brent's method is to combine the bisection
method, the secant method, and inverse quadratic interpolation. Readers
unfamiliar with any of these techniques should refer to CITATIONS. Some
pseudocode is given for Brent's Method in Algorithm~\ref{alg:brent}. The idea
is this:

@@

In general, the inverse quadratic interpolation can proceed very quickly, but we
add the bisection method to ensure convergence. In order to use interpolation,
we require that a few conditions are satisfied. Namely, $s$ has to be in a certain
interval, and the iterates need to obey certain bounds.

The rest of the pseudocode serves only to ensure that variables are assigned
correctly. One notes that $B$ indicates whether the bisection method was used
on the previous iteration.

\begin{algorithm}
  \caption{Using Brent's Method to find a zero of a function}
  \label{alg:brent}
  \begin{algorithmic}
    \Require $f(a_0)f(b_0) > 0$
    \Function{Brent}{$f : [a,b] \to \mathbb{R}$, $a_0$, $b_0$}
      \If{$|f(a_0)| < |f(b_0)|$}
        \Call{swap}{$a_0$,$b_0$}
      \EndIf
      \State $b_{k-1} \gets a_0$
      \State $B \gets$ True \Comment{bisected last time?}
      \While{$f(b_k) \not = 0$ {\bf and} $|b_k - a_k| > \epsilon$}
      \If{$|f(a_k)| < |f(b_k)|$}
      \Call{swap}{$a_k$,$b_k$}
      \EndIf
      \State $b_{k+1} \gets$\Call{inv\_quad\_interp}{$f,a_k,b_k,b_{k-1}$}
      \State $B$ $\gets$ \Call{need\_bisect}{$a_k$,$b_{k-2}$,$b_{k-1}$,$b_k$,$b_{k+1}$,$B$}
      \If{$B$}
        \State $\displaystyle b_{k+1} \gets \frac{a_k+b_k}{2}$ \Comment{bisection method}
      \EndIf
      \State $b_{k-2} \gets b_{k-1}$
      \State $b_{k-1} \gets b_k$
      \If{$f(a_k)f(s) < 0$}~~$b_k \gets s$
      \Else~~$a_k \gets s$
      \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
  \begin{algorithmic}
    \Function{need\_bisect}{$a_k$,$b_{k-2}$,$b_{k-1}$,$b_k$,$b_{k+1}$,$B$}
      \If{$\displaystyle b_{k+1} \not \in \left[ \frac{3a_k+b_k}{4},b_k \right]$}
      \State \Return True
      \ElsIf{$B$ is True}
      \State \Return
        $\displaystyle |b_{k+1}-b_k| \geq \frac{|b_k-b_{k-1}|}{2}$
      \State ~~~~~~~~~~~$\land$~~$|b_k-b_{k-1}| < |\delta|$
      \Else~
      \Return $\displaystyle |b_{k+1}-b_k| \geq \frac{|b_{k-1}-b_{k-2}|}{2}$
      \State ~~~~~~~~~~~~~~$\land$~~$|b_{k-1}-b_{k-2}| < |\delta|$
      \EndIf
    \EndFunction
  \end{algorithmic}
  \begin{algorithmic}
    \Function{inv\_quad\_interp}{$f$,$a$,$b$,$c$}
      \If{$f(a) \not= f(c)$ {\bf and} $f(b) \not= f(c)$}
        \State \Return $\displaystyle \frac{af(b)f(c)}{(f(a)-f(b))(f(a)-f(c))}$
        \State ~~~~~~~~~~~$\displaystyle + \frac{bf(a)f(c)}{(f(b)-f(a))(f(b)-f(c))}$
        \State ~~~~~~~~~~~$\displaystyle + \frac{cf(a)f(b)}{(f(c)-f(a))(f(c)-f(b))}$
      \Else
        \State \Return $\displaystyle b - f(b)\frac{b-a}{f(b)-f(a)}$
      \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

@@

Now that we have developed all of the weight calculations, we can consider using
them to generate a route. Algorithm~\ref{alg:grow} lays out the basic idea in
pseudocode. It omits definitions for many of the functions because they were
covered in sufficient detail in previous sections. The general procedure is
to loop until the algorithm returns to the start point after at least 90\%
of the distance, $d$ is travelled. At each iteration, $\theta$ will be
calculated using \eqref{eq:theta}, where $(x_j,y_j)$ are calculated by applying
Algorithm~\ref{alg:brent} on \eqref{eq:dj}. Once $\theta$ is calculated, a
weight is generated for each node adjacent to $x_i$, and one of the nodes is
selected at random based on their weights.

\begin{algorithm}[t!]
  \caption{Calculating a Route} \label{alg:grow}
  \begin{algorithmic}
    \Require \Call{edges}{} finds adjacent nodes and \Call{random}{} generates
    a random number in the given range.
    \Function{grow\_cycle}{$x_0$,$d$,$e$,$\phi$}
    \State $X \gets \{\}$ \Comment{the route}
    \State $d_k \gets 0$
    \State $\theta \gets \phi$

    \While{$x_{k}\not = x_0$ {\bf or} $d_k < 0.9d$}
      \State $\theta \gets$ \Call{get\_theta}{}

      \State $w \gets 0$ \Comment{sum of all weights}
      \For{$x \in$ \Call{edges}{$x_k$}}
        \State $w \gets w +$\Call{weight}{$x$}
      \EndFor
      \State $q \gets$ \Call{random}{$[0,w]$} \Comment{to choose node}
      \For{$x \in$ \Call{edges}{$x_k$}}
        \If{$q < $ \Call{weight}{$x$,$\theta$}}
          \State $x_{k+1} \gets x$
          \State $d_{k+1} \gets d_k +$\Call{dist}{$x_k$,$x_{k+1}$}
          \State $X \gets X \cup \{x_{k+1}\}$
          \State break
        \Else~$q \gets q -$\Call{weight}{$x$,$\theta$}
        \EndIf
      \EndFor
    \EndWhile
    \State \Return $X$
  \EndFunction

  \end{algorithmic}
\end{algorithm}

@@

\subsection{Distance Calculations}
As mentioned earlier, using the Euclidian norm for latitudes and longitudes
is highly inaccurate; latitudes and longitudes are more similar to 3D polar
coordinates with a fixed $r$ than they are to a coordinate plane. The
haversine formula is used to calculate distances between two points on the
surface of a sphere. The haversine formula uses the haversin function,
which is defined as follows:
\newcommand{\haversin}{\mbox{haversin}}
\[ \haversin (\theta) = \sin^2(\frac{\theta}{2}). \]
Given two coordinate pairs, $(\phi_1, \lambda_1)$ and $(\phi_2, \lambda_2)$,
and a radius, $r$, the haversine formula calculates the distance $d$ between
the two coordinate pairs:
\begin{align*}
  d = r\haversin^{-1}(&\haversin(\phi_2 - \phi_1) + \\
  & \cos(\phi_1)\cos(\phi_2)\haversin(\lambda_2 -\lambda_1)
\end{align*}
Using the definition of the haversin function, this can be reduced into a
function in terms of elementary trigonometric functions. The only non-trivial
thing to note is that $\haversin^{-1}(x) = 2\sin^{-1}(\sqrt{x})$. Armed with
this knowledge, it is fairly easy to reduce. We'll use an intermediate variable
to hold the value under the square root:
\begin{align*}
  h = & \haversin(\phi_2-\phi_1) + \\
  & \cos(\phi_1)\cos(\phi_2)\haversin(\lambda_2-\lambda_1) \\
  = & \sin^2\left(\frac{\phi_2 - \phi_1}{2}\right) + \\
  & \cos(\phi_1)\cos(\phi_2)\sin^2\left(\frac{\lambda_2-\lambda_1}{2}\right)\\
  d = & r\haversin^{-1}\left(\sqrt{h}\right) \\
  = & 2r\sin^{-1}\left(\sqrt{h}\right)
\end{align*}
By simply substituting this as the right hand side of \eqref{eq:dj}, many of the
projection issues can be avoided.

\subsection{Ideal Loop Calculation}

However, the actual ideal loop and, consequently,
\eqref{eq:xj} and \eqref{eq:yj} depend on a set of parametric equations along
a cartesian coordinate plane. These are an inaccurate representation of the
intended ideal loop; as the loop progresses further from the equator, it will
become increasingly squashed in the east/west direction.

To find the proper loop, a cone must be projected from the origin of the sphere,
and the intersection of the cone and the sphere can be used to calculate the
ideal loop. The following equation describes a cone,
\[ 0 = (u \cdot d)^2 - |d||u|\cos \phi , \]
where $u= (x,y,z)$, $d = (a,b,c)$ is a vector parallel to the axis of the
cone, and $\phi$ is one-half the aperture of the cone. Additionally, recall
the equation for a sphere:
\[ r = |u|. \]
Unfortunately, both of these equations are of little aid, since they require
input in the form of cartesian coordinates. We would like to convert these to
a set of equations in spherical form. However, these equations become rather
complicated; for now, we've accepted the projected sphere as ``good enough.''

@@

Unfortunately, the API we used for OSM did not have easy access to elevation
data, something that would have also proved very useful to cyclists. A more
robust implementation of this algorithm would also find the elevation data of
every coordinate and use this information to augment the weights.
